R Tools for Obspack, Footprints and Receptors (rtorf)

### 2024-05-03 README.Rmd
update docs

@@

# TODO ADD identifier of type of altitude of intake_height

@@

```{r}

if(any(names(df) %in% "intake_height")) {
  dfa <- df[,
            max(intake_height),
            by = site_code] |> unique()

  names(dfa)[2] <- "max_altitude"
} else {
  dfa <- df[,
            max(altitude_final),
            by = site_code] |> unique()

  names(dfa)[2] <- "max_altitude"
}
dfa
```

@@

```{r}
df2 <- obs_addtime(df)
```
Then we need a *key_time* to aggregate data.
This can be done using UTC, solar, or local time.
The normal approach is using afteroon solar or local time.

@@

1) solar time
2) local time with columns `site_utc2lst`
3) local time longitude

1) solar time (default)

Here we select the hours of interest and then
aggregate data by year, month and day of solar time.
In this way, we will have one information per day. however
this approach is not appropiate for aircraft
which are aggregated every 10 or 20 seconds. Hence
we need to aggregate data by one time column.
Also, this helps to generate the receptor info 
files including hour, minute and second.
Hence, *we need to add solar or local time column*.

@@

2) local time with column `site_utc2lst`

@@

3) local time longitude

@@

```{r}
df2[, c("solar_time", "solar_time_cut")]
```

@@

```{r}
df3 <- df2[hour_st %in% evening]

@@

```{r}
df4 <- obs_agg(dt = df3,
               gby = "mean",
               cols = c("value",
                        "latitude",
                        "longitude",
                        # "type_altitude",
                        "year_end",
                        "site_utc2lst"),
               verbose = T,
               byalt = TRUE)
```

@@

```{r}
if(any(names(df) %in% "intake_height")) {
  df4[,
      max_altitude := max(intake_height),
      by = site_code]
  df4[,
      c("site_code",
        "intake_height",
        "max_altitude")] |> unique()
} else {
  df4[,
      max_altitude := max(altitude_final),
      by = site_code]
  df4[,
      c("site_code",
        "altitude_final",
        "max_altitude")] |> unique()
}

@@

```{r, eval = F}
readLines(csvy)[1:33]
```

@@

if(any(names(df) %in% "intake_height")) {
receptor <- master[intake_height == max_altitude,
                   c("site_code",
                     "year",
                     "month",
                     "day",
                     "hour", 
                     "minute",
                     "second",
                     "latitude",
                     "longitude",
                     "intake_height",
                     "type_altitude",
                     "year_end",
                     "month_end",
                     "day_end",
                     "hour_end",
                     "minute_end",
                     "second_end")]
receptor$intake_height <- round(receptor$intake_height)

} else {
receptor <- master[altitude_final == max_altitude,
                   c("site_code",
                     "year",
                     "month",
                     "day",
                     "hour",
                     "minute",
                     "second",
                     "latitude",
                     "longitude",
                     "altitude_final",
                     "type_altitude",
                     "year_end",
                     "month_end",
                     "day_end",
                     "hour_end",
                     "minute_end",
                     "second_end")]

receptor$altitude_final <- round(receptor$altitude_final)
}

>>>>

> intake_height or altitude_final

The identification
of the altitude and type is critical. The approach used here 
consists of:

1. Use intake_height
2. Identify `agl` from the name of the tile.
3. If `agl` is not present, `agl = altitude - site_elevation`.
4. If there are some NA in elevation, will result some NA in `agl`
5. A new column is added named `altitude_final` to store `agl` or `asl`
6. Another column named `type_altitude` is added to identify `agl` as 0, or `asl` as 1.

To maintain a coherent approach among text and NetCDF files, 
we will add a column altitude_final and type_altitude. This
is done in `obs_read_nc` or `obs_read`.

@@

@@

```{r}
dfa <- df[,
          max(altitude_final),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
dfa
```

@@

```{r}
df2 <- obs_addtime(df)
```
Then we need a *key_time* to aggregate data.
This can be done using UTC, solar, or local time.
The normal approach is using afternoon solar or local time.

@@

1) Solar time
2) Local time with columns `site_utc2lst`
3) Local time longitude

> solar time (default)

Here we select the hours of interest and then
aggregate data by year, month and day of solar time.
In this way, we will have one information per day. however
this approach is not appropriate for aircraft
which are aggregated every 10 or 20 seconds. Hence
we need to aggregate data by one time column.
Also, this helps to generate the receptor info 
files including hour, minute and second.
Hence, *we need to add solar or local time column*.

@@

> local time with column `site_utc2lst`

@@

> local time longitude

@@

@@

```{r}
df3 <- df2[hour_st %in% evening]
df3[, c("solar_time", "solar_time_cut")]

@@

```{r}
df4 <- obs_agg(dt = df3,
               gby = "mean",
               cols = c("value",
                        "latitude",
                        "longitude",
                        "site_utc2lst"),
               verbose = T,
               byalt = TRUE)
```

@@

```{r}
df4[,
    max_altitude := max(altitude_final),
    by = site_code]
df4[,
    c("site_code",
      "altitude_final",
      "max_altitude")] |> unique()

@@

```{r, eval = F}
readLines(csvy)[1:34] # and more
```

@@

receptor <- master[altitude_final == max_altitude,
                   c("site_code",
                     "year",
                     "month",
                     "day",
                     "hour",
                     "minute",
                     "second",
                     "latitude",
                     "longitude",
                     "altitude_final",
                     "type_altitude",
                     "year_end",
                     "month_end",
                     "day_end",
                     "hour_end",
                     "minute_end",
                     "second_end",
                     "time_decimal")]

receptor$altitude_final <- round(receptor$altitude_final)
